SMT-Boosted Security Types for Low-Level MPC

### 2025-11-18 toplas26/sty.tex
checkpoint

\subsection{Share Typing Example} To illustrate details of the
type system we return to the example of additive sharing in the context
of a 2-client protocol. We first focus on initial secret sharing steps--
let $\prog$ be the following protocol in $\fieldp{2}$ (noting
that $-$ is equivalent to xor in $\fieldp{2}$):
$$
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip)}{1}; \\
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip)}{2} 
\end{array}
$$
We can then assert:
$$
\toeq{\prog} \vdash \varnothing,\varnothing \stredx
\setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
  [\mx{s1}{2}, \elab{\locflip}{1}] : \sharety}, \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
$$
Suppose the protocol goes on to reveal the negation of the
sum (xor-ing) of the inputs:
$$
\begin{array}{lll}
  \px{1} &:=& \elab{\mesg{s2} + \locflip + 1}{1}; \\
  \px{2} &:=& \elab{\mesg{s1} + \locflip}{2};\\
  \out{1} &:=& \elab{\px{1} + \px{2}}{1};\\
  \out{2} &:=& \elab{\px{1} + \px{2}}{2}
\end{array}
$$
Letting this extended protocol be $\prog'$, and letting:
$$
\begin{array}{rcl}
\Gamma &\defeq& \setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
  [\mx{s1}{2}, \elab{\locflip}{1}] : \sharety}\\
R &\defeq&  \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
\end{array}
$$
observe we can then obtain share typings for noninteractive components for
addition of shares by \TirName{Shared} and \TirName{HeAdd}:
$$
\Gamma, \toeq{\prog'} \vdash
[\mx{s2}{1} + \elab{\locflip}{1}, \mx{s1}{2} + \elab{\locflip}{2}] : \sharety
$$
and hence for negation by \TirName{HeAddPub}:
$$
\Gamma, \toeq{\prog'} \vdash
[\mx{s2}{1} + \elab{\locflip}{1} + 1, \mx{s1}{2} + \elab{\locflip}{2}] : \sharety
$$
This implies:
$$
\toeq{\prog'} \vdash \varnothing,\varnothing \stredx
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, R 
$$
Hence by \TirName{PubEntails} we have:
$$
\begin{array}{l}
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, \toeq{\prog'} \vdash
\pubx{\out{1}} : \outty \\
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, \toeq{\prog'} \vdash
\pubx{\out{2}} : \outty
\end{array}
$$

@@

$$
\mems(\Phi) \defeq \{ \store \mid \dom(\store) = \setit{\ x_{\phi}\ \mid\ \phi \in \Phi\ } \}
$$

\begin{definition}
  Given pmf $\pmf$ and $\Phi$ with $\dom(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
  for all $\store \in \mems(\Phi)$ define:
%$$
%\pmf(\phi = v) \defeq \sum_{\store \in \mems(X) \ \text{with}\ \store(\phi) = v} \pmf(\store)
%$$
  $$
  \pmf(\store) \defeq \sum_{\store' \in \mems(X)}
  \begin{cases}
    \pmf(\store') \ \text{if}\ \forall \phi \in \Phi . \store'(\phi) = \store(x_\phi)  \\
    0 \ \text{otherwise}
  \end{cases}
  $$
\end{definition}

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X$ given $\Phi$ where $\bigcup_{\phi \in \Phi} \dom(\phi) \cup X = \dom(\pmf)$ 
  is denoted $\condd{\pmf}{X}{\Phi}$ and defined as follows: 
  $$
  \forall \store_1 \in \mems(X)\ .\ \forall \store_2 \in \mems(\Phi) \ .\ 
  \condd{\pmf}{X}{\Phi}(\store_1 \uplus \store_2) \defeq
  \begin{cases}
    0 \text{\ if\ } \margd{\pmf}{\Phi}(\store_2) = 0\\
    \pmf(\store_1 \uplus \store_2) / \pmf(\store_2) \text{\ otherwise}
  \end{cases}
  $$
\end{definition}

Now we can prove our share typing soundness result, showing that share
typing enforces noninteference modulo output, with our main result
in Theorem \ref{theorem-sty}. We begin by defining a key well-formedness
property of typing components that we show is preserved by non-interactive
typing judgements.

\begin{definition}
  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
  hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $a,b \in \setit{1,2}$ we have
    $
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
    = 
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    $.
  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
    $\toeq{\prog} \models \phi_i^a \eop \rx{w}{\cid_a} \wedge \phi_i^b \eop \phi - \rx{w}{\cid_a}$ for
    some $\phi$ and $\rx{w}{\cid_a}$ then $\rx{w}{\cid_a} \in R$.
  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
  \end{enumerate} 
\end{definition}

\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
    \end{enumerate}
\end{lemma}

\begin{lemma}
  Given $\prog,\Gamma_1,R_1  : {OK}$
  and $\toeq{\prog}  \vdash \Gamma_1, R_1 \stredx \Gamma_2, R_2$.
  Then $\toeq{\prog}, \Gamma_2, R_2 : {OK}$.
\end{lemma}

\begin{definition}
  Given  $\prog,\Gamma,R  : {OK}$ and $\iov(\prog) = S \cup V \cup O$.
  Then $\Gamma$ is \emph{closed for $\prog$} iff for all $H,C$,
  for all $x \in \houtputs$ there exists $\phi$ such that either
  $\Gamma,\toeq{\prog} \vdash [x,\phi] : \sharety$ if $C = \setit{1}$
  or  $\Gamma,\toeq{\prog} \vdash [\phi,x] : \sharety$ if $C = \setit{2}$,
  and for all $\out{\cid} \in O$ we have $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$.
\end{definition}

\begin{theorem}
  \label{theorem-sty}
  If $\prog,\Gamma,R : {OK}$ and $\Gamma$ is closed
  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
\end{theorem}

\begin{theorem}
  \label{theorem-sty}
  If $\toeq{\prog} \vdash \varnothing,\varnothing \stredx \Gamma,R$ and $\Gamma$ is closed
  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
\end{theorem}

>>>>

\subsection{Share Typing Example} To illustrate details of the
type system we return to the example of additive sharing in the context
of a 2-client protocol. We first focus on initial secret sharing steps--
let $\prog$ be the following protocol in $\fieldp{2}$ (noting
that $-$ is equivalent to xor in $\fieldp{2}$):
$$
\begin{array}{lll}
  \elab{\mesg{s1}}{2} := \elab{(\secret{1} \fminus \locflip)}{1}; \quad
  \elab{\mesg{s2}}{1} := \elab{(\secret{2} \fminus \locflip)}{2} 
\end{array}
$$
We can then assert:
$$
\toeq{\prog} \vdash \varnothing,\varnothing \stredx
\setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
  [\elab{\locflip}{1}, \mx{s1}{2}] : \sharety}, \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
$$
Suppose the protocol goes on to reveal the negation of the
sum (xor-ing) of the inputs:
$$
\begin{array}{l}
  \px{1} := \elab{\mesg{s2} + \locflip + 1}{1};\quad \px{2} := \elab{\mesg{s1} + \locflip}{2};\\
  \out{1} := \elab{\px{1} + \px{2}}{1};\quad \out{2} := \elab{\px{1} + \px{2}}{2}
\end{array}
$$
Letting this extended protocol be $\prog'$, and letting:
$$
\begin{array}{rcl}
\Gamma &\defeq& \setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
  [\elab{\locflip}{1}, \mx{s1}{2}] : \sharety}\\
R &\defeq&  \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
\end{array}
$$
observe we can then obtain share typings for noninteractive components for
addition of shares by \TirName{Shared} and \TirName{HeAdd}:
$$
\Gamma, \toeq{\prog'} \vdash
[\mx{s2}{1} + \elab{\locflip}{1}, \mx{s1}{2} + \elab{\locflip}{2}] : \sharety
$$
and hence for negation by \TirName{HeAddPub}:
$$
\Gamma, \toeq{\prog'} \vdash
[\mx{s2}{1} + \elab{\locflip}{1} + 1, \mx{s1}{2} + \elab{\locflip}{2}] : \sharety
$$
This implies:
$$
\toeq{\prog'} \vdash \varnothing,\varnothing \stredx
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, R 
$$
Hence by \TirName{PubEntails} we have:
$$
\begin{array}{l}
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, \toeq{\prog'} \vdash
\pubx{\out{1}} : \outty \\
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, \toeq{\prog'} \vdash
\pubx{\out{2}} : \outty
\end{array}
$$

@@

Share type soundness (Theorem \ref{theorem-sty}) means that share
typing enforces noninterference modulo output. To prove this we need
to assert probabilistic properties in share typings, including for
functions of protocol variables such as reconstruction of shares. So
we make standard extensions of notation for probability distributions
to express probabilities, and conditional probabilities, of functions
of program variables $\phi$. For this extension we posit that
memories can incorporate variables $x_\phi$ that assign valuations
to functions $\phi$.
\begin{definition}
  We let $\Phi$ range over sets of SMT formulas $\phi$, and define
  $
  \mems(\Phi) \defeq \{ \store \mid \dom(\store) = \setit{\ x_{\phi}\ \mid\ \phi \in \Phi\ } \}
  $.
  Given pmf $\pmf$ and $\Phi$ with $\dom(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
  for all $\store \in \mems(\Phi)$ define:
%$$
%\pmf(\phi = v) \defeq \sum_{\store \in \mems(X) \ \text{with}\ \store(\phi) = v} \pmf(\store)
%$$
  $$
  \pmf(\store) \defeq \sum_{\store' \in \mems(X)}
  \begin{cases}
    \pmf(\store') \ \text{if}\ \forall \phi \in \Phi . \store'(\phi) = \store(x_\phi)  \\
    0 \ \text{otherwise}
  \end{cases}
  $$
\end{definition}

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X$ given $\Phi$ where $\bigcup_{\phi \in \Phi} \dom(\phi) \cup X = \dom(\pmf)$ 
  is denoted $\condd{\pmf}{X}{\Phi}$ and defined as follows: 
  $$
  \forall \store_1 \in \mems(X)\ .\ \forall \store_2 \in \mems(\Phi) \ .\ 
  \condd{\pmf}{X}{\Phi}(\store_1 \uplus \store_2) \defeq
  \begin{cases}
    0 \text{\ if\ } \pmf(\store_2) = 0\\
    \pmf(\store_1 \uplus \store_2) / \pmf(\store_2) \text{\ otherwise}
  \end{cases}
  $$
\end{definition}

To proceed, we define a well-formedness property of tuples $\prog,\Gamma,R$
intuitively meaning that all the share types in $\Gamma$ are sound with respect
to $\prog$, and consumed randomness in $R$ as OTPs for initial sharing.
Property (i) ensures that views do not reveal information about secrets.
Property (ii) ensures the key property of sharing-- that individual shares
don't reveal more about secrets than their reconstruction does. Properties
(iii) and (iv) ensure that OTPs used for generating shares or for
secure openings, respectively, are included in $R$.
\begin{definition}
  \label{def-OK}
  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
  hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $a,b \in \setit{1,2}$ we have
    $
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
    = 
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    $.
  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
    $\toeq{\prog} \models \phi_i^a \eop \rx{w}{\cid_a} \wedge \phi_i^b \eop \phi - \rx{w}{\cid_a}$ for
    some $\phi$ and $\rx{w}{\cid_a}$ then $\rx{w}{\cid_a} \in R$.
  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
  \end{enumerate} 
\end{definition}
Next we show that, given well-formed typing environments, share type judgements preserve
the important properties (i-ii) in Definitions \ref{def-OK}. This result follows from
known semi-homomorphic properties of non-interactive operations that are syntactically
verified in typing judgements.
\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
    \end{enumerate}
\end{lemma}
Next, we show that interactive typing rules preserve well-formedness of type
environments. This Lemma follows by Lemma \ref{lemma-sty-noninteractive-sound}
and the security of share generation using OTPs-- that is, shares
generated by $\genshares$ are in uniform marginal distributions. 
\begin{lemma}
  \label{lemma-OK}
  Given $\prog,\Gamma_1,R_1  : {OK}$
  and $\toeq{\prog}  \vdash \Gamma_1, R_1 \stredx \Gamma_2, R_2$.
  Then $\toeq{\prog}, \Gamma_2, R_2 : {OK}$.
\end{lemma}
In addition to typing judgements, in order to show that a protocol is
secure, we need to classify all corrupt views as shares, all
public values as secure openings, and all outputs as reconstructed
shares. We capture this formally with the following notion of
type environment closure.
\begin{definition}
  \label{def-sty-closed}
  Given  $\prog,\Gamma,R  : {OK}$ and $\iov(\prog) = S \cup V \cup O$.
  Then $\Gamma$ is \emph{closed for $\prog$} iff for all $H,C$,
  for all $x \in \houtputs$ there exists $\phi$ such that either
  $\Gamma,\toeq{\prog} \vdash [x,\phi] : \sharety$ if $C = \setit{1}$
  or  $\Gamma,\toeq{\prog} \vdash [\phi,x] : \sharety$ if $C = \setit{2}$,
  and for all $\out{\cid} \in O$ we have $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$.
\end{definition}
Now, the following is a consequence of Definitions \ref{def-OK} and \ref{def-sty-closed}, and then
our main Theorem \ref{theorem-sty} follows by Lemma \ref{lemma-OK} and Theorem \ref{theorem-OK}.
\begin{theorem}
  \label{theorem-OK}
  If $\prog,\Gamma,R : {OK}$ and $\Gamma$ is closed
  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
\end{theorem}
\begin{theorem}
  \label{theorem-sty}
  If $\toeq{\prog} \vdash \varnothing,\varnothing \stredx \Gamma,R$ and $\Gamma$ is closed
  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
\end{theorem}
